---
title: "Asn2 Q2"
author: "Sumanth Donthula"
date: "2022-09-20"
output: pdf_document
---

2.a.

The values of Beta0 and Beta1 are 2.114,0.038

The model is #Y=2.11+0.03883X

```{r}
Data =read.delim("AS_2_Q_2_Data.txt", header = FALSE, sep = " ")
Y=Data[,2]
X=Data[,6]

Model=lm(Y~X)

Beta0=as.numeric(Model$coefficients["(Intercept)"])
Beta0
Beta1=as.numeric(Model$coefficients["X"])
Beta1
```

2.b.

The plot of the Data and the Model Fit is as shown. The regression function fit the model well with minimum outliers.

```{r}
plot(X,Y,xlab="ACT Test Score",ylab="Average GPA")
abline(Beta0,Beta1,col=2,lwd=2)
```
2.c.

The point estimate of mean freshman GPA of ACT score=30 is 3.2788

```{r}
Y1=Beta0+Beta1*30
Y1
```
2.d.

The point estimate of change in mean response when entrance score is increased 
by one point is Beta1 which is 0.0388

Problem 3

3.a.

The 95 % confidence interval of test score which is 28 is (3.06,3.34)

```{r}
xnew = data.frame(X=28)

Y2=Beta0+Beta1*28

SXY = sum(X *Y ) - length(X) * mean(X) * mean(Y)
SYY = sum( Y * Y) - length(Y) * (mean(Y))^2
SXX = sum( X * X) - length(X) * (mean(X))^2
RSS = SYY - Beta1 * SXY
SIGMA=(RSS/(length(X)-2))^0.5
mu=mean(X)

n=0.05
t=qt(n/2, length(X)-2, lower.tail=TRUE)
#Method 1 from Function
confInt = predict(Model,xnew, interval = "confidence", level = 0.95, se.fit = TRUE)
conflwr=confInt$fit[2]
confupr=confInt$fit[3]
conflwr
confupr
#Method 2 from Formula
ConfUb=Y2+t*SIGMA*((1/length(X))+((xnew-mu)^2/SXX))^0.5
ConfLb=Y2-t*SIGMA*((1/length(X))+((xnew-mu)^2/SXX))^0.5
ConfUb
ConfLb
```
3.b.

The 95 % prediction interval of test score which is 28 is (1.95,4.44)

```{r}
#Method 1 from Function
predInt=predict(Model,xnew, interval = "prediction", level = 0.95, se.fit = TRUE)
predlwr=predInt$fit[2]
predlwr
predupr=predInt$fit[3]
predupr
#Method 2 Derivating with Formula
PredUb=Y2+t*SIGMA*(1+(1/length(X))+((xnew-mu)^2/SXX))^0.5
PredLb=Y2-t*SIGMA*(1+(1/length(X))+((xnew-mu)^2/SXX))^0.5
PredUb
PredLb
```
3.c.

Yes the Prediction interval is wider than confidence interval. Because the prediction model is used to find estimates of the random samples rather than the confidence interval which is an inference on sample data.

3.d.

The range of confidence band is (3.02,3.37).

Yes the confidence band is little wider because it represents the entire regression model line not only the sample at Xh=28.

```{r}
Weight = sqrt(2*qf(0.95,2,length(X)-2))

cbandupper = confInt$fit[,1]+Weight*confInt$se.fit
cbandlower = confInt$fit[,1]-Weight*confInt$se.fit
cbandlower
cbandupper
```
4.a.

Anova Table

```{r}
Anova =anova(Model)
Anova
```

4.b.

MSR is the sum of squares due to regression taking degrees of freedom into account.

MSE is Mean Squared Error which is not a biased estimator of Standard Deviation
(Variance Squared).

When Beta1=0 or the slope of the regression equation is 0.

4.c.

Let's Start Null Hypothesis Testing

Assumption:
H0: Beta1 is zero
Beta1=0

Ha: Beta1 is not Zero
Beta1!=0

Decision rule:
The value of Fscore is less than FStarScore so we reject the null hypothesis so the Alternative is true which is Beta1 is not equal to 0.

Conclusion:
Beta1 is not Zero
Beta1!=0

```{r}
alpha=0.01
Fscore = qf((1-alpha),1,length(X)-2)
Fscore

FStarScore=9.242 #From Anova Table
FStarScore
```
4.d.

The absolute magnitude of reduction in variance of Y when X is introduced in
regression model is R Squared(R^2)

4.e.

The sign of R is positive as the data has a positive correlation from the graph.

```{r}
R=sqrt(0.07262) # from summary
R
```
4.f.

R Squared gives clear cut interpretation. It describes the percent of variance 
of Y with respect to X. R square is usually used to represent the relation. It
takes values from 0 to 1.

